Using reference models in feature selection \vspace{.1in}

### 2019-12-06 paper/ref_approach.tex
update bodyfat plots
edits

\begin{document}
\maketitle
\begin{abstract}
  Feature selection, or more generally, model reduction is an important aspect of the statistical workflow aiming to provide insights from data. In this paper we discuss and demonstrate the benefits of using a reference model in feature selection. A reference model acts as a noise-filter on the target variable by modeling its data generating mechanism. As a result, using the reference model predictions in the model selection procedure reduces the variability and improves stability leading to improved model selection performance. Assuming that a Bayesian reference model describes the true distribution of future data well, the theoretically preferred usage of the reference model is to project its predictive distribution to a reduced model leading to projection predictive variable selection approach. Alternatively, reference models may also be used in an ad-hoc manner in combination with common feature selection methods.  In several numerical experiments, we demonstrate superior performance of projection predictive approach and show how use of a reference model translates into better and more stable feature selection independently of the specific selection procedure. 
% These benefits are also illustrated on a real-world example of predicting the amount of body fat for which we compare a reference model based selection method to a purely data driven approach.
\end{abstract}

@@

In statistical applications, one of the main steps in the modelling
workflow is covariate or feature selection, which is a special case of
model reduction. Variable selection may have multiple goals.  First,
if the covariates themselves are of interest, we can use variable
selection to infer which variables contain predictive information
about the target variable.  Second, as simpler models come with the
advantages of reduced measurement costs and improved interpretability,
we may be interested in finding the minimal subset of covariates which
still provides good predictive performance (or good balance between
simplicity and predictive performance).  When the predictive
capability is guiding the selection, the true data generation
mechanism of future data can be approximated either by using the
observed data directly or alternatively by using predictions from a
\emph{reference model} \citep{vehtari2012survey}.

In data based approaches, such as Lasso selection
\citep{tibshirani1996regression} or stepwise backward/forward
regression \citep{venables2013modern,harrell2015regression}, the
observed empirical data distribution is utilised as a proxy of future
data usually in combination with cross-validation or information
criteria to provide estimates for out-of-sample predictive
performance.  In contrast, reference model based methods approximate
the future data generation mechanism using the predictive distribution
of a reference model, which can be, for example, a full-encompassing
model including all covariates.  The main assumption under any
reference model approach is that we operate in an
$\mathcal{M}$-\textit{complete} framework
\citep{book:bernardo_smith,vehtari2012survey}, that is, we have
constructed a model which reflects our beliefs about
the future data in the best possible way and which has passed model checking and criticism
\citep[see, e.g.][]{gelman2013bayesian,gabry2019visualization}.  The reference model approach
has been used in Bayesian statistics at least since the seminal
work of \citet{paper:reference_lindley}. For more historical
references, see the papers by \citet{vehtari2012survey} and
\citet{paper:model_selection}, and for most recent methodological
development see \citet{paper:projpred}.

@@

In this section, we will provide an initial motivation and intuition 
for the use of reference models to improve feature selection procedures.
We will start with a case study that is repeatedly used throughout
the paper to illustrate the benefits of reference models before we dive deeper
into the theoretical reasons why reference models help in feature selection.

@@

The target variable of interest is the amount of body fat, which is
measured by immersing a person in water tank which is a complex and
expensive procedure. Additionally, we have information about 13
covariates which are anthropometric measurements (e.g., height, weight
and thicknes of skin and fat in different body locations). The
covariates are are highly correlated which causes additional
challenge in the feature selection. In total, we have 251
observations. The goal is to find the model which is able to predict
the amount of fat well while requiring least amount of
measurements for a new person.

@@

For the selection via projpred, the Bayesian reference model includes
all the covariates using a regularised horseshoe prior~\citep{paper:rhs} on the
covariate coefficients. Submodels are explored using
forward search (the results are not sensitive to whether forward or
backward search is used), and the predictive utility is the expected
log-predictive density (elpd) estimated using approximate leave-one-out cross-validation via Pareto-smoothed importance-sampling
\citep[PSIS-LOO-CV; ][]{paper:psis_loo}.  We select the smallest  submodel with an elpd score similar to the reference model
when taking into account the uncertainty in estimating the predictive
model performance. See Appendix A for brief review of projection
predictive approach and \citet{paper:model_selection} and
\citet{paper:projpred} for more details. The complete projpred procedure is implemented in the \texttt{projpred} R package.

The inclusion frequencies of each feature in the final model given 100
bootstrap samples are depicted in Figure
\ref{fig:inclusion_frequencies}. The projpred approach has only two
variables, `abdomen' and `weight', with inclusion frequencies above
50\% (`abdomen' is the only one included always), the third most
included is `wrist' at 44\%, and the fourth is `height' at 35\%, while
steplm has seven features above 50\%. Such a higher variability and
lower stability of steplm can be observed also in the bootstrap model
selection frequencies reported in Table
\ref{tab:model_frequencies}. For example, the first five selected
models have a cumulative frequency of 76\% with projpred, but only of
14\% with steplm. In addition, the size of the selected model is much
smaller when using projpred.
\begin{figure}[tp]
  \centering
  \includegraphics[width=0.98\textwidth]{graphics/inc_prob.pdf}
  \caption{Bootstrap inclusion frequencies after 100 bootstrap samples.}
  \label{fig:inclusion_frequencies}
\end{figure}

\begin{table}[tp]
\footnotesize
\centering
\begin{tabular}{l|lr|lr}
%  \hline
M & projpred & Freq \% & steplm & Freq \%  \\ 
  \hline
1 & abdom, weight & 39 & abdom, age, forearm, height, hip, neck, thigh, wrist & 4 \\
2 & abdom, wrist & 10 & abdom, age, chest, forearm, height, neck, thigh, wrist & 4 \\
3 & abdom, height & 10 & abdom, forearm, height, neck, wrist & 2 \\
4 & abdom, height, wrist & 9 & abdom, forearm, neck, weight, wrist & 2 \\
5 & abdom, weight, wrist & 8 & abdom, age, height, hip, thigh, wrist & 2 \\
6 & abdom, chest, height, wrist & 2 & abdom, age, height, hip, neck, thigh, wrist & 2 \\
7 & abdom, biceps, weight, wrist & 2 & abdom, age, ankle, forearm, height, hip, neck, thigh, wrist & 2 \\
8 & abdom, height, weight, wrist & 2 & abdom, age, biceps, chest, height, neck, wrist & 2 \\
9 & abdom, age, wrist & 2 & abdom, age, biceps, chest, forearm, height, neck, thigh, wrist & 2 \\
10 & abdom, age, height, neck, thigh, wrist & 2 & abdom, age, ankle, biceps, weight, wrist & 2 \\
%   \hline
\end{tabular}
\caption{Bootstrap model selection frequencies after 100 bootstrap samples.}
\label{tab:model_frequencies}
\end{table}

In the first two rows of Table \ref{tab:model_performances}, we report the
predictive performances, in terms of cross-validated root mean square
error (RMSE), of the full model and the selected models using either
the projection predictive approach or the stepwise backward
elimination.  There is no difference in predictive performance of the
selected models by different approaches even if there is difference in
the selected features.

We repeat the experiment with a modified data set by adding 84
unrelated noisy features. In the last two rows of Table
\ref{tab:model_performances}, we report the cross-validated RMSE, the size of
the selected model and the number of selected noisy features using
either the projection predictive approach or the stepwise backward
elimination.  The projection predictive approach has similar
predictive performance and the same submodel size as in the first
experiment, whereas the stepwise regression has worse predictive
performance and selects a much larger submodel including a large
number of irrelevant features.

\begin{table}[tp]
\footnotesize
\centering
\begin{tabular}{ll|rrrr}
%  \hline
Data  & Method & RMSE Full & RMSE Selected & \# Selected & \# Selected noisy \\ 
  \hline
original & projpred & 4.35 & 4.47 & 2 &   \\
& steplm & 4.38 & 4.47 & 7 &  \\
\hline
+ noisy features & projpred & 4.58 & 4.54 & 2 & 0  \\
& steplm & 5.89 & 5.43 & 19 & 13 \\
%   \hline
\end{tabular}
\caption{Model predictive performances with original data (first two
  rows) and with extra noisy features (last two rows) estimated with
  10-fold cross-validation. Abbreviations: Full = full model; Selected =
  selected submodel; \# Selected = total number of features selected;
  \# Selected noisy = number of extra noisy features selected.}
\label{tab:model_performances}
\end{table}

@@

If one of the models to be compared is the full model which can be
used as a reference model, there is no additional cost of using a
reference model. Sometimes including all the available features in an
elaborate model can often be computationally demanding.  In such a
case, even simpler screening or dimensionality reduction techniques,
as for example the supervised principal components
\citep{bair2006prediction,piironen2018} can produce useful reference
models.

@@

A good predictive model is able to filter part of the noise present in
the data. The noise is the main source of the instability and tends to
obscure the relevance of the covariates in relation to the target
variable of interest. Let us consider the following simple
explanatory example taken from \cite{paper:projpred}. The data
generation mechanism is
\begin{alignat}{2} \label{eq:simulated_data}
     &f\sim \N(0,1) && \nonumber \\ 
     Y|&f\sim \N(f,1) && \\
     X_{j}|&f \overset{iid}{\sim} \N(\sqrt{\rho}f,1-\rho) \quad &&j=1,\ldots,k \nonumber \\
     X_{j}|&f \overset{iid}{\sim} \N(0,1) &&j=k+1,\ldots,p \nonumber,
\end{alignat}
where $f$ is the latent variable of interest of which $Y$ is a noisy
observation. The first $k$ covariates are strongly related to the
target variable $Y$ and correlated among themselves. Precisely, $\rho$
is the correlation among any pair of the first $k$ covariates, whereas
$\sqrt{\rho}$ and $\sqrt{\rho/2}$ are the level of correlation between
any relevant covariate and, respectively, $f$ and $Y$. If we had an
infinite amount of observations, the sample correlation would be equal
to the true correlation between $X_j$ and $Y$. However, even in this
ideal asymptotic regime, this correlation would still remain a biased
indicator of the true relevance of each covariate (represented by the
correlation between $X_j$ and $f$) due to the intrinsic noisy nature
of $Y$.

@@

\begin{figure}[tp]
  \centering
  \vspace{-5mm}
  \includegraphics[width=0.4\textwidth]{graphics/correlation.pdf}
  \caption{Sample correlation plot of each feature (relevant in red, 
  non-relevant in blue) with the target variable $y$ and the latent variable 
  $f$ respectively on the x and y axis. Data are generated according to 
  procedure \eqref{eq:simulated_data} with parameters $n=70$, $\rho=0.3$,
  $k=100$, $p=1000$.\\}
  \label{fig:correlation}
\end{figure}

@@

In the bodyfat example above, we desired at the same time good
predictive performance and selection of a smaller number of
features. When the goal is to select a minimal subset of features,
which have similar predictive performance as the full model, we call it
\emph{minimal subset feature selection}. This minimal subset might
exclude features which have some predictive information about the
target, but given the minimal subset these features are not able to
provide such additional information that would improve predictive
performance. The usual reason for this is that the relevant features
not in the minimal are subset highly correlated with features already in
the minimal subset. We will return to a problem of finding all
the features with some predictive power in Section~\ref{complete-selection}.

@@

When the variable selection is repeated with different simulated data
sets, there is some variability in which variables are selected. We
measure the stability of variable selection by computing the entropy
of observed distribution over different models. The smallest entropy
would be obtained if the approach always selects the same set of
variables, and the largest entropy would be observed if the approach
would always select different set of variables.
Figure \ref{fig:entropy} reports the entropy scores for the different compared
methods. Lower entropy corresponds to a more stable
selection. Highly correlated predictive features may happen to be
alternately selected, making the measure of the stability of the
selection a non-trivial task. Entropy cannot distinguish the interchangeability 
due to correlation from instability. Thus, such a measure should be 
considerd as a relative and not as an absolute measure of stability.
The use of a reference model
improves slightly the stability of steplm in feature selection. The
projection predictive approach results to be far more stable than all
other approaches. This is likely due to projpred being based on better
decision theoretical formulation which 1) takes into account the full
predictive distribution and not just point estimate and 2) projects
the reference model posterior to the submodel instead of using a
simple refit of submodels.

@@

Here we repeat the selection of Section \ref{intro-example} via
stepwise backward regression. In this case, the overall number of
covariates (original plus noisy) is 100, as it was in the last part of
Section \ref{intro-example}. We compare results with and without using
a reference model on top of the procedure following a simple
reference model approach outlined in Equation \eqref{eq:ref_mod} 
of Section \ref{simulations}. Figure
\ref{fig:bodyfat_step_refvsdata} shows the number of irrelevant
features included in the final model and the out-of-sample root mean
square error (RMSE). Results are based on 100 bootstrap samples on the
whole dataset, and the predictive performance is tested on the
observations excluded at each bootstrap sample. We observe that the
reference model reduces the number of irrelevant features included in
the final model. This leads to less overfitting and thus to improved
out-of-sample predictive performance in terms of RMSE. The reference
model approach applied to the stepwise backward regression achieves
outstanding improvements considering its simplicity, yet it does not
reach the goodness of the much more sophisticated projective
prediction approach (see results of Section \ref{intro-example}).
\begin{figure}[tp]
  \centering
  \includegraphics[width=0.98\textwidth]{graphics/bodyfat_step_refvsdata.pdf}
  \caption{Stepwise backward selection with and without using a reference model. The x-axis denotes the number of selected irrelevant features on the left and the out-of-sample RMSE on the right-hand side based on 100 bootstrap samples.\\}
  \label{fig:bodyfat_step_refvsdata}
\end{figure}

@@

An alternative to minimal subset feature selection is \emph{complete
  feature selection} in which the goal is to find all relevant
features that have some predictive information about the target. In
complete feature selection, it is possible that there are theoretically
relevant features, but given finite noisy data we are not able infer their relevance.
The projection predictive approach was originally designed for the minimal subset
feature selection, but we generalize it here for the complete
feature selection case. In addition, we analyse the benefits of the reference
model approach in combination with three other complete feature
selection approaches.

As criteria of selection performance, we evaluate the average false
discovery rate and the average
sensitivity (i.e., the ratio of the number of relevant selected
features over the total number of relevant features). We also provide
a comparison of the stability of the selection by means of a stability
measure proposed by \citet{paper:stability}.

@@

We can use the projection predictive approach for complete feature
selection by using it iteratively.  Applying the straightforward
implementation of projpred, we are able to select a minimal subset of
features, which yield to a model with a predictive performance
comparable to the full model one.  The iterative projection repeats
the projpred selection for different iterations, at each time
excluding from the search the features selected in the previous
iterations. At each iteration the selected submodel size corresponds to the
one having a predictive performance close enough to the baseline
model, which in this iterative version is the submodel with the
highest predictive score explored at the current iteration. The
stopping rule that we adopted is the following:
\begin{equation} 
\label{eq:rule_of_thumb}
\text{min} \{i\in \{0,\ldots,p\}: \text{P}(\text{elpd}_{i}-\text{elpd}_{\text{best}}>0)\geq \alpha \},
\end{equation}
where $i$ indexes the submodel size and ``best'' stays for the best predictive explored submodel
at the considered iteration. The algorithm terminates when the empty model (only intercept) satisfies
the stopping rule (see Algorithm \ref{alg:iterated_proj}).
The choice of the hyperparameter $\alpha$ can be non-trivial, we have 
observed sensitivity of the selection to such a choice mainly when using cross-validation with a small
number of observations or not very predictive features. 
In our experiments we chose the default value used in the \texttt{projpred} 
\texttt{R}-package that corresponds to $\alpha=0.16$.

@@

\subsection{Alternative complete feature selection approaches}

@@

We consider three alternative complete selection procedures,
 the control of the local
false discovery rate \citep{paper:efron, efron2012large}, the
empirical Bayes median \citep{johnstone2004needles}, and the selection
by posterior credible intervals.

@@

The empirical Bayes median procedure consists of fitting a Bayesian
model with a prior composed by a mixture of a delta spike in zero and
a heavy-tailed distribution. We use the implementation in the
\texttt{R}-package \texttt{EbayesThresh}
\citep{Silverman+etal:2017:EbayesThresh}. As suggested by
\cite{johnstone2004needles}, we use a Laplace distribution resulting
in a thresholding property, that is, there exists a threshold value
such that all the data under that threshold have posterior median
equal to zero. Therefore the selection is done by selecting only those
parameters whose posterior median is different from zero. The
hyperparameter of the Laplace distribution and the mixing weight of
the prior are estimated by marginal maximum likelihood.

@@

All of these methods provide a complete selection procedure and we
compare their performance with and without using a reference model.
That is, in the data condition, we apply the method on the original
data $y$ while, in the referece model condition, we replace $y$ by
their mean predictions $\hat{y}$ based on the reference model.

@@

To investigate the performance of the four complete selection procedures,
we are going to use simulations based on the normal means problem.
The normal means problem consists of estimating the (usually
sparse) vector of means of a vector of normally distributed
observations. The dimensionality of the vector of means is denoted by
$p$ and $\{z_{j}\}_{j=1}^{p}$ is the vector of observations of the
random variables $\{Z_{j}\}_{j=1}^{p}$. The task is to estimate the
latent variables $\{\theta_{j}\}_{j=1}^{p}$ of the following model: \
\begin{equation}\label{eq:normal_means_problem}
Z_{j}|\theta_{j},\sigma\overset{ind}{\sim}\N(\theta_{j},\sigma^{2}), \quad j=1,\ldots,p.
\end{equation}
This is equivalent to a linear regression where the design matrix is
the identity matrix with the number of features being equal to the
number of observations. This formulation can be found in practice, for
example, in the analysis of microarray data, where a large set of
genes are tested in two groups of patients labeled as positive or
negative to some disease \citep{paper:efron,efron2012large}. The
objective of the analysis is to select the subset of genes
statistically relevant to the disease. One common way to proceed is to
compute the two-sample $t$-statistic for every gene separately.  After
normalising these statistics, they then become the data $Z$ in the
normal means problem \eqref{eq:normal_means_problem}. For further
details see the examples by \cite{paper:efron, efron2012large}.

>>>>

\begin{document}
\maketitle
\begin{abstract}
  Feature selection, or more generally, model reduction is an important aspect of the statistical workflow aiming to provide insights from data. In this paper we discuss and demonstrate the benefits of using a reference model in feature selection. A reference model acts as a noise-filter on the target variable by modeling its data generating mechanism. As a result, using the reference model predictions in the model selection procedure reduces the variability and improves stability leading to improved model selection performance. Assuming that a Bayesian reference model describes the true distribution of future data well, the theoretically preferred usage of the reference model is to project its predictive distribution to a reduced model leading to projection predictive variable selection approach. Alternatively, reference models may also be used in an ad-hoc manner in combination with common feature selection methods.  In several numerical experiments, we demonstrate superior performance of projection predictive approach and show how use of a reference model translates into better and more stable feature selection independently of the specific selection method. 
% These benefits are also illustrated on a real-world example of predicting the amount of body fat for which we compare a reference model based selection method to a purely data driven approach.
\end{abstract}

@@

In statistical applications, one of the main steps in the modelling
workflow is covariate or feature selection, which is a special case of
model reduction. Feature (variable) selection may have multiple goals.  First,
if the covariates themselves are of interest, we can use variable
selection to infer which variables contain predictive information
about the target variable.  Second, as simpler models come with the
advantages of reduced measurement costs and improved interpretability,
we may be interested in finding the minimal subset of covariates which
still provides good predictive performance (or good balance between
simplicity and predictive performance).  When the predictive
capability is guiding the selection, the true data generation
mechanism of future data can be approximated either by using the
observed data directly or alternatively by using predictions from a
\emph{reference model} \citep{vehtari2012survey}.

In data based approaches, such as Lasso selection
\citep{tibshirani1996regression} or stepwise backward/forward
regression \citep{venables2013modern,harrell2015regression}, the
observed empirical data distribution is utilised as a proxy of future
data usually in combination with cross-validation or information
criteria to provide estimates for out-of-sample predictive
performance.  In contrast, reference model based methods approximate
the future data generation mechanism using the predictive distribution
of a reference model, which can be, for example, a full-encompassing
model including all covariates.  The main assumption under any
reference model approach is that we operate in an
$\mathcal{M}$-\textit{complete} framework
\citep{book:bernardo_smith,vehtari2012survey}, that is, we have
constructed a model which reflects our beliefs about
the future data in the best possible way and which has passed model checking and criticism
\citep[see, e.g.][]{gelman2013bayesian,gabry2019visualization}.  The reference model approach
has been used in Bayesian statistics at least since the seminal
work of \citet{paper:reference_lindley}. For more historical
references, see the papers by \citet{vehtari2012survey} and
\citet{paper:model_selection}, and for most recent methodological
development see a paper by \citet{paper:projpred}.

@@

In this section, we will provide an initial motivation and intuition 
for the use of reference models to improve feature selection methods.
We will start with a case study that is repeatedly used throughout
the paper to illustrate the benefits of reference models before we dive deeper
into the theoretical reasons why reference models help in feature selection.

@@

The target variable of interest is the amount of body fat, which is
measured by immersing a person in water tank which is a complex and
expensive procedure. Additionally, we have information about 13
covariates which are anthropometric measurements (e.g., height, weight
and circumference of different body parts). The
covariates are highly correlated which causes additional
challenge in the feature selection. In total, we have 251
observations. The goal is to find the model which is able to predict
the amount of body fat well while requiring least amount of
measurements for a new person.

@@

For the selection via projpred, the Bayesian reference model includes
all the covariates using a regularised horseshoe prior~\citep{paper:rhs} on the
covariate coefficients. Submodels are explored using
forward search (the results are not sensitive to whether forward or
backward search is used), and the predictive utility is the expected
log-predictive density (elpd) estimated using approximate leave-one-out cross-validation via Pareto-smoothed importance-sampling
\citep[PSIS-LOO-CV; ][]{paper:psis_loo}.  We select the smallest  submodel with an elpd score similar to the reference model
when taking into account the uncertainty in estimating the predictive
model performance. See Appendix A for brief review of projection
predictive approach, and papers by \citet{paper:model_selection} and
\citet{paper:projpred} for more details. The complete projpred approach is implemented in the \texttt{projpred} R package.

The inclusion frequencies of each feature in the final model given 100
bootstrap samples are shown in Figure
\ref{fig:inclusion_frequencies}. In case of projpred there are two
variables, `abdomen' and `weight', which have inclusion frequencies
above 50\% (`abdomen' is the only one included always), the third most
frequently included is `wrist' at 44\%, and the fourth one is `height'
at 35\%.  The steplm approach has seven features with inclusion
frequencies above 50\%. Such a higher variability and lower stability
of steplm can be observed also in the bootstrap model selection
frequencies reported in Table \ref{tab:model_frequencies}. For
example, the first five selected models have a cumulative frequency of
76\% with projpred, but only of 14\% with steplm. In addition, the
sizes of the selected models with projpred are much smaller with
steplm.
\begin{figure}[tp]
  \centering
  \includegraphics[width=0.98\textwidth]{graphics/inc_prob.pdf}
  \caption{Bootstrap inclusion frequencies after 100 bootstrap samples.}
  \label{fig:inclusion_frequencies}
\end{figure}

\begin{table}[tp]
\footnotesize
\centering
\begin{tabular}{l|lr|lr}
%  \hline
M & projpred & Freq \% & steplm & Freq \%  \\ 
  \hline
1 & abdom., weight & 39 & abdom., age, forearm, height, hip, neck, thigh, wrist & 4 \\
2 & abdom., wrist & 10 & abdom., age, chest, forearm, height, neck, thigh, wrist & 4 \\
3 & abdom., height & 10 & abdom., forearm, height, neck, wrist & 2 \\
4 & abdom., height, wrist & 9 & abdom., forearm, neck, weight, wrist & 2 \\
5 & abdom., weight, wrist & 8 & abdom., age, height, hip, thigh, wrist & 2 \\
6 & abdom., chest, height, wrist & 2 & abdom., age, height, hip, neck, thigh, wrist & 2 \\
7 & abdom., biceps, weight, wrist & 2 & abdom., age, ankle, forearm, height, hip, neck, thigh, wrist & 2 \\
8 & abdom., height, weight, wrist & 2 & abdom., age, biceps, chest, height, neck, wrist & 2 \\
9 & abdom., age, wrist & 2 & abdom., age, biceps, chest, forearm, height, neck, thigh, wrist & 2 \\
10 & abdom., age, height, neck, thigh, wrist & 2 & abdom., age, ankle, biceps, weight, wrist & 2 \\
%   \hline
\end{tabular}
\caption{Bootstrap model selection frequencies after 100 bootstrap samples.}
\label{tab:model_frequencies}
\end{table}

The first two rows of Table \ref{tab:model_performances} show the
predictive performances, in terms of cross-validated root mean square
error (RMSE), of the full model and the selected models using projpred
or steplm.  There is no signfiant difference in predictive performance
of the selected models by different approaches even if there is
difference in the selected features. This can be explained by high
correlation between the features and different combinations can
provide similar predictive accuracy.

We repeat the experiment with a modified data set by adding 84
unrelated noisy features. The last two rows of Table
\ref{tab:model_performances} show the cross-validated RMSE, the size
of the selected model and the number of selected noisy features using
projpred or steplm.  The projection predictive approach has similar
predictive performance and the same number of selected features as
with the original data, whereas the stepwise regression has worse
predictive performance and the number of selected features is much
higher and include a large number of irrelevant features.

\begin{table}[tp]
\footnotesize
\centering
\begin{tabular}{ll|rrrr}
%  \hline
Data  & Method & RMSE Full & RMSE Selected & \# Selected & \# Selected noisy \\ 
  \hline
bodyfat & projpred & 4.35 & 4.47 & 2 &   \\
& steplm & 4.38 & 4.47 & 7 &  \\
\hline
+ noisy features & projpred & 4.58 & 4.54 & 2 & 0  \\
& steplm & 5.89 & 5.43 & 19 & 13 \\
%   \hline
\end{tabular}
\caption{Predictive performances with original bodyfat data (first two
  rows) and with extra noisy features (last two rows) estimated with
  10-fold cross-validation. Abbreviations: Full = full model; Selected =
  selected submodel; \# Selected = selected number of features;
  \# Selected noisy = selected number of extra noisy features.}
\label{tab:model_performances}
\end{table}

@@

If one of the models to be compared is the full model which can be
used as a reference model, there is no additional cost of using a
reference model. Sometimes including all the available features in an
elaborate model can be computationally demanding.  In such a
case, even simpler screening or dimensionality reduction techniques,
as for example the supervised principal components
\citep{bair2006prediction,piironen2018} can produce useful reference
models.

@@

A good predictive model is able to filter part of the noise present in
the data. The noise is the main source of the instability and tends to
obscure the relevance of the covariates in relation to the target
variable of interest. We demonstrate with the following simple
explanatory example taken from \cite{paper:projpred}. The data
generation mechanism is
\begin{alignat}{2} \label{eq:simulated_data}
     &f\sim \N(0,1) && \nonumber \\ 
     Y|&f\sim \N(f,1) && \\
     X_{j}|&f \overset{iid}{\sim} \N(\sqrt{\rho}f,1-\rho) \quad &&j=1,\ldots,k \nonumber \\
     X_{j}|&f \overset{iid}{\sim} \N(0,1) &&j=k+1,\ldots,p \nonumber,
\end{alignat}
where $f$ is the latent variable of interest of which $Y$ is a noisy
observation. The first $k$ covariates are strongly related to the
target variable $Y$ and correlated among themselves. Precisely, $\rho$
is the correlation among any pair of the first $k$ covariates, whereas
$\sqrt{\rho}$ and $\sqrt{\rho/2}$ are the level of correlation between
any relevant covariate and, respectively, $f$ and $Y$. If we had an
infinite amount of observations, the sample correlation would be equal
to the true correlation between $X_j$ and $Y$. However, even in this
ideal asymptotic regime, this correlation would still remain a biased
indicator of the true relevance of each covariate (represented by the
correlation between $X_j$ and $f$) due to the intrinsic noisy nature
of $Y$.

@@

\begin{figure}[tp]
  \centering
  \vspace{-5mm}
  \includegraphics[width=0.4\textwidth]{graphics/correlation.pdf}
  \caption{Sample correlation plot of each feature (relevant in red, 
  non-relevant in blue) with the target variable $y$ and the latent variable 
  $f$ respectively on the x and y axis. Data are generated according to 
  \eqref{eq:simulated_data} with parameters $n=70$, $\rho=0.3$,
  $k=100$, $p=1000$.\\}
  \label{fig:correlation}
\end{figure}

@@

In the bodyfat example above, we desired at the same time good
predictive performance and selection of a smaller number of
features. When the goal is to select a minimal subset of features,
which have similar predictive performance as the full model, we call it
\emph{minimal subset feature selection}. This minimal subset might
exclude features which have some predictive information about the
target, but given the minimal subset these features are not able to
provide such additional information that would improve predictive
performance significantly. The usual reason for this is that the relevant features
which are not in the minimal subset are highly correlated with features already in
the minimal subset. We will return to a problem of finding all
the features with some predictive power in Section~\ref{complete-selection}.

@@

When the variable selection is repeated with different simulated data
sets, there is some variability in selected variables. We
measure the stability of variable selection by computing the entropy
of observed distribution over different models. The smallest entropy
would be obtained if the approach always selects the same set of
variables, and the largest entropy would be observed if the approach
would always select different set of variables.
Figure \ref{fig:entropy} shows the entropy scores for the different compared
methods. Lower entropy corresponds to a more stable
selection. Highly correlated predictive features may happen to be
alternately selected, making the measure of the stability of the
selection a non-trivial task. Entropy cannot distinguish the interchangeability 
due to correlation from instability. Thus, such a measure should be 
considerd as a relative and not as an absolute measure of stability.
The use of a reference model
improves slightly the stability of steplm in feature selection. The
projection predictive approach results to be far more stable than all
other approaches. This is likely due to projpred being based on better
decision theoretical formulation which 1) takes into account the full
predictive distribution and not just point estimate and 2) projects
the reference model posterior to the submodel instead of using a
simple refit of submodels.

@@

Here we repeat the selection of Section \ref{intro-example} via
stepwise backward regression. In this case, the overall number of
covariates (original plus noisy) is 100, as it was in the last part of
Section \ref{intro-example}. We compare results with and without using
a simple reference model approach outlined in \eqref{eq:ref_mod} with
steplm. Figure \ref{fig:bodyfat_step_refvsdata} shows the number of
irrelevant features included in the final model and the out-of-sample
root mean square error (RMSE). Results are based on 100 bootstrap
samples on the whole dataset, and the predictive performance is tested
on the observations excluded at each bootstrap sample. We observe that
the reference model reduces the number of irrelevant features included
in the final model. This leads to less overfitting and thus to
improved out-of-sample predictive performance in terms of RMSE. The
reference model approach applied to the stepwise backward regression
achieves outstanding improvements considering its simplicity, yet it
does not reach the goodness of the much more sophisticated projective
prediction approach (see results of Section \ref{intro-example}).
\begin{figure}[tp]
  \centering
  \includegraphics[width=0.98\textwidth]{graphics/bodyfat_step_refvsdata.pdf}
  \caption{Stepwise backward selection with and without using a reference model. The x-axis denotes the number of selected irrelevant features on the left and the out-of-sample RMSE on the right-hand side based on 100 bootstrap samples.\\}
  \label{fig:bodyfat_step_refvsdata}
\end{figure}

@@

An alternative to minimal subset feature selection is \emph{complete
  feature selection} in which the goal is to find all relevant
features that have some predictive information about the target. In
complete feature selection, it is possible that there are theoretically
relevant features, but given finite noisy data we are not able infer their relevance.
The projection predictive approach was originally designed for the minimal subset
feature selection, but we test here a simple iterative variant for the complete
feature selection case. In addition, we analyse the benefits of the reference
model approach in combination with three other
methods which have been originally designed for complete feature selection.

As the criteria of selection performance, we evaluate the average false
discovery rate and the average
sensitivity (i.e., the ratio of the number of relevant selected
features over the total number of relevant features). We also provide
a comparison of the stability of the selection by means of a stability
measure proposed by \citet{paper:stability}.

@@

We modify the projection predictive approach for complete feature
selection by using it iteratively.  Applying the straightforward
implementation of projpred, we are able to select a minimal subset of
features, which yield to a model with a predictive performance
comparable to the full model one.  The iterative projection repeats
the projpred selection for different iterations, at each time
excluding from the search the features selected in the previous
iterations. At each iteration the selected submodel size corresponds to the
one having a predictive performance close enough to the baseline
model, which in this iterative version is the submodel with the
highest predictive score explored at the current iteration. The
stopping rule that we adopted is the following:
\begin{equation} 
\label{eq:rule_of_thumb}
\text{min} \{i\in \{0,\ldots,p\}: \text{P}(\text{elpd}_{i}-\text{elpd}_{\text{best}}>0)\geq \alpha \},
\end{equation}
where $i$ indexes the submodel size and ``best'' stays for the best predictive explored submodel
at the considered iteration. The algorithm terminates when the empty model (only intercept) satisfies
the stopping rule (see Algorithm \ref{alg:iterated_proj}).
The choice of the hyperparameter $\alpha$ is non-trivial, and we have 
observed sensitivity of the selection to such a choice mainly when using cross-validation with a small
number of observations or not very predictive features. 
In our experiments we chose the default value used in the \texttt{projpred} 
\texttt{R}-package that corresponds to $\alpha=0.16$.

@@

\subsection{Alternative complete feature selection methods}

@@

We consider three alternative complete selection methods,
 the control of the local
false discovery rate \citep{paper:efron, efron2012large}, the
empirical Bayes median \citep{johnstone2004needles}, and the selection
by posterior credible intervals.

@@

The empirical Bayes median approach consists of fitting a Bayesian
model with a prior composed by a mixture of a delta spike in zero and
a heavy-tailed distribution. We use the implementation in the
\texttt{R}-package \texttt{EbayesThresh}
\citep{Silverman+etal:2017:EbayesThresh}. As suggested by
\cite{johnstone2004needles}, we use a Laplace distribution resulting
in a thresholding property, that is, there exists a threshold value
such that all the data under that threshold have posterior median
equal to zero. Therefore the selection is done by selecting only those
parameters whose posterior median is different from zero. The
hyperparameter of the Laplace distribution and the mixing weight of
the prior are estimated by marginal maximum likelihood.

@@

All of these methods provide a complete selection approach and we
compare their performance with and without using a reference model.
That is, in the data condition, we apply the method on the original
data $y$ while, in the referece model condition, we replace $y$ by
their mean predictions $\hat{y}$ based on the reference model.

@@

To investigate the performance of the four complete selection approaches,
we are going to use simulations based on the normal means problem.
The normal means problem consists of estimating the (usually
sparse) vector of means of a vector of normally distributed
observations. The dimensionality of the vector of means is denoted by
$p$ and $\{z_{j}\}_{j=1}^{p}$ is the vector of observations of the
random variables $\{Z_{j}\}_{j=1}^{p}$. The task is to estimate the
latent variables $\{\theta_{j}\}_{j=1}^{p}$ of the following model: \
\begin{equation}\label{eq:normal_means_problem}
Z_{j}|\theta_{j},\sigma\overset{ind}{\sim}\N(\theta_{j},\sigma^{2}), \quad j=1,\ldots,p.
\end{equation}
This is equivalent to a linear regression where the design matrix is
the identity matrix with the number of features being equal to the
number of observations. This formulation can be found in practice, for
example, in the analysis of microarray data, where a large set of
genes are tested in two groups of patients labeled as positive or
negative to some disease \citep{paper:efron,efron2012large}. The
objective of the analysis is to select the subset of genes
statistically relevant to the disease. One common way to proceed is to
compute the two-sample $t$-statistic for every gene separately.  After
normalising these statistics, they then become the data $Z$ in the
normal means problem \eqref{eq:normal_means_problem}. For further
details see the examples by \cite{paper:efron, efron2012large}.
